---
title: "TEND_Group_Testing"
author: "Me"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

## R Markdown

### Setting Working Directory

```{r setup, include=TRUE, echo=TRUE}
# Trying here(), which will also set working directory based off your Rmd file
library(here)
here::i_am("TEND_Group_Testing.Rmd") # tells you full file path script's current location
here() # returns file path of where the script is current saved
setwd(here()) # sets the working directory to be wherever your source file or Rmd is

getwd() # why not double check?
```

### Adding libraries/packages

```{r prep, include=T, echo=T}
library(tidyverse) 
# Loads 'ggplot2' for data visualization. 'dplyr' for data manipulation, 'tidyr' for 
# data tidying, 'readr' for data import, 'purrr' for functional programming, 'tibble' 
# for (tibbles) a modern re-imagining of data frames, 'stringr' for strings
library(writexl)
# Allows yoyu to save an explored data fram as an Excel file. It will go to your wd.
library(foreign) 
# Read data stored by SPSS and Stata
library(psych) 
# For personality, psychometric, and psychological research, includes describe function 
# and error bars
library(stargazer) 
# Handy regression tables
library(readr)
```

### Load and Prepare Data

```{r load in, include=FALSE}
Brain_Data <- read.csv("Brain_Data.csv", header=T, sep=",", na.strings=c("NA", "888", "999"))
Behavioral_Data<-read.csv("Bx_Data.csv", header=T, sep=",", na.strings=c("NA", "888", "999"))

# Merge the data sets
DATA <- merge(Brain_Data, Behavioral_Data, "ID", all=T)
View(DATA)

# The most common form of data is factors, numbers, strings and characters. They are all 
# understood by R in different ways and certain functions need data to be in specific types 
# to be understood correctly. Below are examples of rewriting the variable "ID" as a 
# factor. And also creating a new variables Age as the numeric version of "Age.at.V1". 
# To call a variable within a data set you say [name of dataset]$[name of varible], 
# if the name of the variable does not exist, it will CREATE the variable in that data 
# set. So in this case, "Age" did not exist, it was just Age.at.V1 and Age.at.V2 in our 
# sheet, so instead we created a number version of "Age.at.V1" now called "Age"
DATA$ID<- as.factor(DATA$ID)
DATA$Age<- as.numeric(DATA$Age_at_Bx)

# It will be a very good habit to not only check what form your data or columns are 
# represented as, it will be very useful to learn the functions that change your code 
# into a more effective form. 
# 
# attach() and detach()
# If you have several variables you are manipulating and don't want to call the data 
# frame they belong to each time ([data frame]$[varibale you want]) you can "attach" 
# the data set so that R assumes all variables you call until you "detach" it will be 
# in the data frame you are looking at. Below is an example of scoring a measurement 
# of several components. Instead of having to write "DATA$RADS_DM" etc. for each subscale 
# we are adding, we can just write the variable name. THIS WAS COPY+PASTED FROM 
# JOHANNA's Rmd
attach(DATA)
DATA$RADS_total_rescore<- RADS_DM + RADS_AN + RADS_NS + RADS_SC
detach(DATA)
View(DATA)
# This would be useful in a situation where you are constantly having to call a column 
# that is nested within a variable, or multiple. This could get confusing when you have 
# performed significant exploration and have many variables in your environment

# attach() and detach()
# If you have several variables you are manipulating and don't want to call the data 
# frame they belong to each time, you can use attach() and detach() to tell the software 
# to call only to the data frame you specify. This prevents you from having to type out 
# the long code that calls to a specific column nested in a data set
attach(DATA)
DATA$RADS_total_rescore<- RADS_DM + RADS_AN + RADS_NS + RADS_SC
detach(DATA)

# recode() COPY+PASTED FROM JOHANNA'S Rmd
# #Here is how you can easily recode something using the "recode" function, the first 
# value will be what the data currently is, and the second value is what you now want 
# it to be. For example, we want to round up all of the scores in one column or variable, 
# 'Puberty_rounded'.
DATA$Puberty_rounded <- recode(DATA$Puberty, '1.5' = 2.0, '2.5' = 3.0, '3.5' = 4.0, 
                               '4.5' = 5.0)
# Dplyr has some very useful functions as well, but this is useful to know

# creating new variable and column
# This is how you can add a new variable with NA values
DATA$attempt_reason<- NA

# assigning NA to conditions in a column
# By using the variable assignment operator '<-', we can assign NA avlues to a specified 
# column in a df, and add conditions so that only subjects with scores below 21, for 
# example, get reassigned to NA
DATA$Bx_to_Scan_Days[DATA$Bx_to_Scan_Days > 21] <- NA

# rename()
# If you wanted to rename variables or names of columns, you can change them with 
# rename(), even doing multiple rows simultaneously
DATA_Final<- rename(DATA,"Hx_Attempt" = "Hx.Attempt", "Age_First.Attempt" = "Age.First.Attempt","Number_of_Attempts" = "Number.of.Attempts")

# Let's now save our new data sheet using the writexl library
library(writexl)
write_xlsx(DATA_Final, "Example_DATA_Export.xlsx")
```

### Quick Check!
###### Johanna gave a very useful tip for quickly checking your data (certain columns, variables) before running anlyses or continuing with your exploratory analysis. Using View() and data.frame() in conjuction with one another gives you a quick glimpse at a quickly-made df of your specified columns.

```{r quick check, include=TRUE, echo=TRUE}
# This is a great way of double-checking that your preparation hasn't changed the data. 
# Think of it as a way of looking twice before turning!
View(data.frame(DATA_Final$ID, DATA_Final$CDRSR_total, DATA_Final$Sex, DATA_Final$Puberty))
```

### Start Exploring

```{r explore, include=TRUE, echo=TRUE}

# First we will use colnames() to see the names of the variables of our data set
# Assign the colnames to a variable through a data frame, to be accessed later saves 
# this as a data frame
Column_indexes <- data.frame(colnames(DATA))
str(Column_indexes)

# which() allows us to know the position of a specific variable, ex "Gender" is our 
# 10th column variable
which(colnames(DATA) == "Gender")

# rowMeans()
# We will now use rowMeans() to get the average of multiple variables. rowMeans can 
# also have the 'na.rm' logical argument, which, when TRUE, will go through your data 
# and ignore any rows that have NAs in your specified columns.As you can 
# see, we assign a new column with the '<-' and '$' 
DATA$RADS_average<- rowMeans(DATA[,c("RADS_DM",	"RADS_AN",	"RADS_NS",	"RADS_SC")], 
                             na.rm=TRUE)

# WIDE vs LONG format CONVERSION , common in longitudinal studies/data

# Long format will mean there are several rows per participant that are differentiated 
# by time point. Wide is when you have one row per participant and the repeated measures 
# and different column. Certain functions require the longitudinal data to be in wide or 
# long format.
# Here is an example of a simple reshape from long to wide where time points already 
# designated. For the sake of an example, let's say we called for View() after reading 
# in our data, and we realize that our data is in long format and it needs to be in wide.
## CCTG_Data_Long <- read.csv("fake_longitudinal_data.csv", header=T, sep=",", 
##     na.strings=c("NA", "888", "999"))

## CCTG_Data_Wide <- reshape(CCTG_Data_Long, idvar = "pid", timevar = "visit_cycle", 
##     direction = "wide")
```

# FORLOOPS !!!
```{r forloop ex, include=F, echo=T, warning=FALSE}
# Complex example of reshape where "instances" of a measure need to be assigned with 
# for loop!

# Let's break down each point in this long function.Let's say our problem is that we 
# are trying to convert long to wide, but we don't have a 'marker' or way of assigning 
# our data to a certain visit, whether it be their first or third. In order to convert 
# to wide format, the computer needs to know

# reading in data
## Contact_Log_Long <- read.csv("fake_contact_log.csv", header=T, sep=",", 
##                             na.strings=c("NA", "888", "999"))
## turning Contact_Num into missing values
## Contact_Log_Long$Contact_Num <- NA
## for(x in 1:length(Contact_Log_Long$pid)){
##   if (x == 1)
##     {Contact_Log_Long$Contact_Num[x] = 1}
##  else if (Contact_Log_Long$pid[x] == Contact_Log_Long$pid[x-1]) 
##  {Contact_Log_Long$Contact_Num[x] = Contact_Log_Long$Contact_Num[x-1] + 1}
##  else if (Contact_Log_Long$pid[x] != Contact_Log_Long$pid[x-1]) 
##  {Contact_Log_Long$Contact_Num[x] = 1}
##  }

## MI_Data_Wide <- reshape(Contact_Log_Long, idvar = "pid", timevar = "Contact_Num", 
##     direction = "wide")
```

### forloop example
##### Due to the importance of understanding forloops beyond the input-output, I will perform a similar function to the one that Johanna made, but on a different data frame either made by me or found online (or...both?)
```{r forloop contd, include=F, echo=FALSE}
# First, we create the necessary variables to construct the long format data frame. 
# 'subID' is a vector that represents exactly what you think, subjects!
subID <- c(1, 1, 1, 2, 2, 2, 3, 3, 3)
# 'eventID' is a vector that represents exactly what you think, events!
eventID <- c("Event1", "Event2", "Event3")
# Random values for each event
ran_values <- rnorm(9)  

long_practicdf <- data.frame(SubjectID = rep(subject_ids, each = 3),
                      EventID = rep(event_ids, times = 3),
                      Value = values)
print(long_df)

# Create a function to add visit number based on subject ID
add_visit_number <- function(data) {
  unique_subjects <- unique(data$SubjectID)
  for (subject in unique_subjects) {
    subject_rows <- data$SubjectID == subject
    data$VisitNumber[subject_rows] <- 1:length(data$VisitNumber[subject_rows])
  }
  return(data)
}

# Add visit number using the function
long_df <- add_visit_number(long_df)
print(long_df)

# Convert to wide format using reshape()
wide_df <- reshape(long_df, idvar = "SubjectID", timevar = "EventID", direction = "wide")
print(wide_df)

```

### mutate()

```{r mutate, include=T, echo=T}
# coming from dplyr, the primary purpose of mutate() is to add new columns to a data 
# frame based on calculations or transformations applied to existing columns. It allows 
# you to perform various operations, such as mathematical calculations, conditional 
# operations, string manipulations, and more. LEARN THIS FUNCTION
```

### DURING-Session REFLECTION

##### 

### POST-Session REFLECTION

##### TBD..
